On Misuses of the Kolmogorov--Smirnov Test for One-Sample Goodness-of-Fit

### 2023-01-10 manuscript/manuscript.tex
bootstrap steps for section 4

The invalidity of the statndard statistic for serially dependent data can be
illustrated by a simple simulation study. Consider a first-order autoregressive
or AR(1) model with marginal standard normal distribution. For each AR
coefficient $\psi \in \{-0.3, 0.2, 0.1, 0, 0.1, 0.2, 0.3\}$, we generated 1000
series of length 100. For each series, we applied the KS statistic
in~\eqref{eq:ks_standard} to test $H_0$ that the $X_i$'s follow $N(0, 1)$
marginally. The histograms of the p-values from $1000$ replicates for all the
$\psi$ values are displayed in Figure~\ref{fig:hist_correlation}. When
$\phi \ne 0$, the histograms do not look like to be generated from a standard
uniform distribution. Positive $\psi$ values led to liberal p-values while
negative $\psi$ values led to conservative p-values. The larger the magnitude of
$\psi$, the bigger the deviation from the desired level.

@@

The parameters of the working normal copula are
estimated from fitting an auto-regressive moving average (ARMA) model to the
observed data transformed to the standard normal scale. In particular, let
$Z_i = \Phi\{ F(X_i)\}$, $i = 1, \ldots, n$, where $\Phi$ is the distribution
function of the standard normal. Then we fit an ARMA$(p, q)$ model to
$Z_1, \ldots, Z_n$ with AR order $p$ and MA order $q$ selected by the Akaike
Information Criterion (AIC). This can be done with, for example, function
\texttt{auto.arima()} from  R package \texttt{forecast}
\citep{hyndman2008automatic}. Since it is known that the mean of $Z_i$'s are
zero, it is necessary to set the intercept of the ARMA model as zero in the
fitting process. The selected ARMA$(p, q)$ model with fitted
parameters will be used as the working model to generate bootstrap sample with
serial dependence mimicking that in the observed data. The unconditional
variance $\sigma^2$ of the ARMA model can be obtained with function
\texttt{tacvfARMA()} from R package \texttt{ltsa} \citep{mcleod2007algorithms}.

@@

The procedure demonstrated in Section~\ref{sec:dependence} works when the data
is dependent and the hypothesized distribution is completely specified. However,
this is not practical. In practice we do not know the parameters of the
hypothesized distribution. Therefore, it is valuable to have a procedure that
corrects for both fitted parameters and serially dependent data. The remedy
using copulas can be modified to be effective in the case where both assumptions
must be violated. The null hypothesis is equivalent to Section~\ref{sec:fitted}
and the test statistic is equal to Equation~\eqref{eq:ks_fitted}. The bootstrap
procedure in Section~\ref{sec:fitted} is no longer valid because the serial
dependence is not accounted for. Let $r$ be the AR(1) coefficient of the normal
copula that matches the sample first-order Spearman's rho of the observed data
as obtained by the \texttt{iRho} function in the \textsf{copula} package
\citep{Rpkg:copula}. We propose the following bootstrap procedure to assess the
significance of the observed KS statistic.

\begin{enumerate}
\item
  Generate $Z_1, \ldots, Z_n$ from an AR(1) process with autocorrelation
  coefficient $r$ such that the $Z_i$'s are $N(0, 1)$ variables.
\item
  Form a bootstrap sample $X_i^* = F^{-1}_{\hat\theta_n} [\Phi(Z_i)]$,
  $i = 1, \ldots, n$, whose first-order sample Spearman's rho matches that of
  the observed data.
\item
  Fit $F_\theta$ to the sample $X_1^*, \ldots, X_n^*$ and obtain estimator
  $\hat\theta_n^*$
\item
  Obtain the empirical distribution function $F_n^*$ of the bootstrap sample
  $X_1^*, \ldots, X_n^*$.
\item
  Calculate bootstrap KS statistic
  \[
    D^* = \sup_x \lvert F_n^* (x)- F_{\hat\theta_n^*}(x) \rvert.
  \]
\item
  Repeat the previous steps a large number $B$ times and use the empirical
    distribution of the $B$ test statistics to approximate
    the null distribution of the observed statistic.
\end{enumerate}

>>>>

The invalidity of the standard statistic for serially dependent data can be
illustrated by a simple simulation study. Consider a first-order autoregressive
or AR(1) model with marginal standard normal distribution. For each AR
coefficient $\psi \in \{-0.3, 0.2, 0.1, 0, 0.1, 0.2, 0.3\}$, we generated 1000
series of length 100. For each series, we applied the KS statistic
in~\eqref{eq:ks_standard} to test $H_0$ that the $X_i$'s follow $N(0, 1)$
marginally. The histograms of the p-values from $1000$ replicates for all the
$\psi$ values are displayed in Figure~\ref{fig:hist_correlation}. When
$\phi \ne 0$, the histograms do not look like to be generated from a standard
uniform distribution. Positive $\psi$ values led to liberal p-values while
negative $\psi$ values led to conservative p-values. The larger the magnitude of
$\psi$, the bigger the deviation from the desired level.

@@

The parameters of the working normal copula are
estimated from fitting an auto-regressive moving average (ARMA) model to the
observed data transformed to the standard normal scale. In particular, let
$Z_i = \Phi^{-1}\{ F(X_i)\}$, $i = 1, \ldots, n$, where $\Phi$ is the distribution
function of the standard normal. Then we fit an ARMA$(p, q)$ model to
$Z_1, \ldots, Z_n$ with AR order $p$ and MA order $q$ selected by the Akaike
Information Criterion (AIC). This can be done with, for example, function
\texttt{auto.arima()} from  R package \texttt{forecast}
\citep{hyndman2008automatic}. Since it is known that the mean of $Z_i$'s are
zero, it is necessary to set the intercept of the ARMA model as zero in the
fitting process. The selected ARMA$(p, q)$ model with fitted
parameters will be used as the working model to generate bootstrap sample with
serial dependence mimicking that in the observed data. The unconditional
variance $\sigma^2$ of the ARMA model can be obtained with function
\texttt{tacvfARMA()} from R package \texttt{ltsa} \citep{mcleod2007algorithms}.

@@

For testing distributions with unspecified parameters with serially dependent
data, we can modify the procedure in the last section. In preparation, let
$\hat\theta_n$ be the marginally fitted parameters, and let
$Z_i = \Phi^{-1} \{F_{\hat\theta_n}(X_i)\}$, $i = 1, \ldots, n$.
For $Z_i$'s that are marginally standard normal variables, we fit an
ARMA$(p,q)$ model with AR order $p$ and MA order $q$ selected by AIC as in the
last section. Let $\sigma^2$ be the unconditional variance of the fitted ARMA
process. We propose the following bootstrap procedure to assess the
significance of the observed KS statistic.

\begin{enumerate}
\item
  Generate $Z_1^*, \ldots, Z_n^*$ from the fitted ARMA$(p, q)$ process
  with innovation variance $1 / \sigma^2$ such that the $Z_i$'s are $N(0, 1)$ variables.
\item
  Form a bootstrap sample $X_i^* = F^{-1}_{\hat\theta_n} [\Phi(Z_i)]$,
  $i = 1, \ldots, n$, whose first-order sample Spearman's rho matches that of
  the observed data.
\item
  Fit $F_\theta$ to the sample $X_1^*, \ldots, X_n^*$ and obtain estimator
  $\hat\theta_n^*$
\item
  Obtain the empirical distribution function $F_n^*$ of the bootstrap sample
  $X_1^*, \ldots, X_n^*$.
\item
  Calculate bootstrap KS statistic
  \[
    D^* = \sup_x \lvert F_n^* (x)- F_{\hat\theta_n^*}(x) \rvert.
  \]
\item
  Repeat the previous steps a large number $B$ times and use the empirical
    distribution of the $B$ test statistics to approximate
    the null distribution of the observed statistic.
\end{enumerate}
